---
title: "Explore and Summarize Red Wine Data"
author: "Luis Cruz"
date: "10 Feb 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

In this project a wine dataset is used.
It is related to red variant of the Portuguese "Vinho Verde" wine.
The data is available [here](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityReds.csv).
A description of this dataset can be found [here](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt).

According to this description there are no Null values.
A description of the attributes is also given:

   1 - **fixed acidity** (tartaric acid - g / dm^3): most acids involved with wine or fixed or nonvolatile (do not evaporate readily)

   2 - **volatile acidity** (acetic acid - g / dm^3): the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste

   3 - **citric acid** (g / dm^3): found in small quantities, citric acid can add 'freshness' and flavor to wines

   4 - **residual sugar** (g / dm^3): the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet

   5 - **chlorides** (sodium chloride - g / dm^3): the amount of salt in the wine

   6 - **free sulfur dioxide** (mg / dm^3): the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine

   7 - **total sulfur dioxide** (mg / dm^3): amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

   8 - **density** (g / cm^3): the density of water is close to that of water depending on the percent alcohol and sugar content

   9 - **pH**: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

   10 - **sulphates** (potassium sulphate - g / dm3): a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant

   11 - **alcohol** (% by volume): the percent alcohol content of the wine

   12 - **quality** (score between 0 and 10) -- __Output variable (based on sensory data)__
   

## Univariate Plots Section

```{r read}
path <- "./wineQualityReds.csv"
#path <- "https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityReds.csv"
df <- read.csv(path)
```

### Overview
```{r overview}
dim(df)
names(df)
str(df)
summary(df)
```

This overview shows that all the attributes are numerical. Also we can see that the wines in this study have relatively low alcohol, since 75% of wines have between 8.40 and 11.10 by volume. This is a common feature in this kind of wine, "Vinho Verde".
Properties such as __residual.sugar__, __chlorides__, __free.sulfur.dioxide__ and total.sulfur.dioxide seem to have outliers since the maximum is much higher than the 75% percentile.

### Alcohol
```{r acidity}
qplot(data = df, x=alcohol, binwidth=0.5)+
  scale_x_continuous(limits = c(8,15), breaks=seq(8,15,0.5))

qplot(data = df, x=alcohol, binwidth=0.1)+
  scale_x_continuous(limits = c(8,15), breaks=seq(8,15,0.5))
```

There are some outliers below 9% and above 14% alcohol by volume. Most wines are between 9% and 12%.

### pH
```{r pH}
qplot(data = df, x=pH, binwidth=0.1)+
  scale_x_continuous(limits = c(2.7,4.1), breaks=seq(2.7,4.1,0.1))
```